\section*{Discussion}

Building on Shannon's formulation of information theory, there have been
significant efforts using this theoretical framework to understand the
information processing capabilities of biological systems, and the evolutionary
consequences for organisms harboring signal transduction systems
\cite{Bergstrom2004, Taylor2007, Tkacik2008, Polani2009, Nemenman2010,
Rivoire2011}. Recently, with the mechanistic dissection of molecular signaling
pathways, significant progress has been made on the question of the physical
limits of cellular detection and the role that features such as feedback loops
play in this task \cite{Bialek2005, Libby2007, Tkacik2011, Rhee2012a,
Voliotis2014a}. But the field still lacks a rigorous experimental test of these
ideas with precision measurements on a system that is tractable both
experimentally and theoretically.

In this paper we take advantage of the recent progress on the quantitative
modeling of input-output functions of genetic circuits to build a minimal model
of the simple repression motif \cite{Phillips2019}. By combining a series of
studies on this circuit spanning diverse experimental methods for measuring
gene expression under a myriad of different conditions, we possess complete a
priori parametric knowledge -- allowing us to generate parameter-free
predictions for processes related to information processing. Some of the model
parameters for our kinetic formulation of the input-output function are
informed by inferences made from equilibrium models. We use the fact that if
both, kinetic and thermodynamic languages describe the same system, the
predictions must be self-consistent. In other words, if the equilibrium model
can only make statements about the mean mRNA and mean protein copy number
because of the way these models are constructed, those predictions must be
equivalent to what the kinetic model has to say about these same quantities.
This condition therefore constrains the values that the kinetic rates in the
model can take. To test whether or not the equilibrium picture can reproduce
the predictions made by the kinetic model we compare the experimental and
theoretical fold-change in protein copy number for a suite of biophysical
parameters and environmental conditions (\fref{fig3_cell_cycle}(C) upper row).
The agreement between theory and experiment demonstrates that these two
frameworks can indeed make consistent predictions.

The kinetic treatment of the system brings with it increasing predictive power
compared to the equilibrium picture. Under the kinetic formulation, the
predictions are not limited only to the mean but to any of the moments of the
mRNA and protein distributions. We first test these novel predictions by
comparing the noise in protein copy number (standard deviation / mean) with
experimental data. Our minimal predicts the noise up to a systematic deviation.
The physical or biological origins of this discrepancy remains an open
question. In that way the work presented here exposes the status quo of our
understanding of gene regulation in bacteria, posing new questions to be
answered with future refinements of the model. We extended our analysis to
infer entire protein distributions at different input signal concentrations by
using the maximum entropy principle. What this means is that we compute moments
of the protein distribution, and then use these moments to build an
approximation to the full distribution. These predicted distributions are then
compared with experimental single-cell distributions as shown in
\fref{fig4_maxent}(B) and \siref{supp_maxent}. Again here although our minimal
model systematically underestimates the width of the distributions by a
constant multiplicative factor, it serves to inform how to expect changes in
parameters such as protein copy number or protein-DNA binding affinity will
affect the full probabilistic input-output function of the genetic circuit up
to a multiplicative constant. We then use our model to predict the information
processing capacity.

By maximizing the mutual information between input signal concentration and
output protein distribution over all possible input distributions we predict
the channel capacity of the system over a suite of biophysical parameters such
as varying repressor protein copy number and repressor-DNA binding affinity.
Although there is no reason to assume the the simplified synthetic circuit we
used as an experimental model operates optimally given the distribution of
inputs, the relevance of the channel capacity comes from its interpretation as
a metric of the physical limits of how precise the inference that cells can
make about what the state of the environment is. Our model, despite the
systematic deviations, makes non-trivial predictions such as the existence of
an optimal repressor copy number for a given repressor-DNA binding energy,
predicting the channel capacity up to an additive constant (See
\fref{fig5_channcap}). We note that this differs from previous theoretical
results since this optimal combination does not come from adding a cost term
for the regulation \cite{Tkacik2011}. This is a consequence of the parameters
inferred in \cite{Razo-Mejia2018} for the allosteric repressor never allowing
all repressors to go into the inactive (non-DNA binding) state. That means that
even at saturating concentrations of inducer, as the number of repressors
increases, a significant number of them are still able to bind to the promoter.
This causes all of the input-output functions to be biased towards low
expression levels, decreasing the amount of information that the circuit is
able to process.

As we have done throughout this work, we consider it important to highlight the
limitations of the work presented here. The previously discussed systematic
deviation for the noise and skewness of the predicted distributions (See
\siref{supp_multi_gene}), and therefore of the predicted distributions and
channel capacity remains an unresolved question that deserves to be addressed
in further iterations of our minimal model. Also, as first reported in
\cite{Razo-Mejia2018}, our model fails to capture the steepness of the
fold-change induction curve for the weakest repressor binding site (See
\fref{fig3_cell_cycle}(B)). Furthermore the minimal model in
\fref{fig2_minimal_model}(A), despite being widely used, is an
oversimplification of the physical picture of how the transcriptional machinery
works. The coarse-graining of all the kinetic steps involved in the
transcription initiation into two effective promoter states -- active and
inactive -- ignores potential kinetic regulatory mechanisms of intermediate
states \cite{Scholes2017}. Moreover it has been argued that despite the fact
that the mRNA count distribution does not follow a Poisson distribution, this
effect could be caused by unknown factors not at the level of transcriptional
regulation \cite{Choubey2018}.

The findings of this work open the opportunity to accurately test intriguing
ideas that connect Shannon's metric of how accurately a signaling system can
infer the state of the environment, with Darwinian fitness \cite{Taylor2007}.
Beautiful work along these lines has been done in the context of the
developmental program of the early {\it Drosophila} embryo \cite{Tkacik2008,
Petkova2019}. These studies demonstrated that the input-output function of the
pair-rule genes works at channel capacity, suggesting that selection has acted
on these signaling pathways, pushing them to operate at the limit of what the
physics of these systems allows. Our system differs from the early embryo in
the sense that we have a tunable circuit with variable amounts of information
processing capabilities. Furthermore, compared with the fly embryo in which the
organism tunes both the input and output distributions over evolutionary time,
we have experimental control of the distribution of inputs that the cells are
exposed to. Consequently this means that instead of seeing the final result of
the evolutionary process, we can set different environmental challenges, and
track over time the evolution of the population. These experiments could shed
light into the suggestive hypothesis of information bits as a trait on which
natural selection acts. We see this exciting direction as part of the overall
effort in quantitative biology of predicting evolution \cite{Lassig2017}.